\documentclass[a4,11pt]{article}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}

%% \title{Analysis of the {\ttfamily GSOEP9402} data}
%% \author{Arthur Allignol}
%% \date{}

\usepackage{booktabs}
\usepackage[osf]{sourcesanspro}
\usepackage{sourcecodepro}
\usepackage[T1]{fontenc}
%\renewcommand*{\ttdefault}{sourcecodepro}
\usepackage{subcaption}
\usepackage{graphicx,rotating,epsfig,multirow,multicol,hhline}
\usepackage{amsmath,amsthm,amssymb,amsfonts}
\usepackage{url}
\usepackage{hyperref}

\renewcommand{\familydefault}{\sfdefault}
\usepackage{sfmath}


\def\bea{\begin{eqnarray*}}
\def\eea{\end{eqnarray*}}
%Dense format
\setlength{\parindent}{0em} \setlength{\textwidth}{16cm} \setlength{\textheight}{25cm}
\setlength{\topmargin}{-0.5cm} \setlength{\oddsidemargin}{0cm} \setlength{\headheight}{0cm}
\setlength{\headsep}{0cm}
\newcommand{\nat}{{\it I\hspace{-0.12cm}N}}
\pagestyle{empty}


\newcommand{\p}{{\rm P}}
\newcommand{\erw}{{\rm E}}
\newcommand{\var}{{\rm var}}
\newcommand{\eins}{{\bf 1}}%{\mathbf{1}}
\newcommand{\dif}{{\rm d}}
\newcommand{\cif}{{\rm CIF}}
\newcommand{\Prob}{\mathbb{P}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\D}{\mathrm{d}}

%Special commands:
\newcommand{\Bin}{\operatorname{Bin}} % Binomial Distribution
\newcommand{\NegBin}{\operatorname{NegBin}} % Negative Bin
\newcommand{\HypGeom}{\operatorname{HypGeom}} % Hypergeometric Distribution
\newcommand{\Pois}{\operatorname{Po}} % Hypergeometric Distribution
\newcommand{\Po}{\operatorname{Po}} %
\newcommand{\Exp}{\operatorname{Exp}} %
\newcommand{\Par}{\operatorname{Par}} %
\newcommand{\Ga}{\mathcal{G}a} %
\newcommand{\Be}{\mathcal{B}e} %
\newcommand{\Var}{\operatorname{Var}} %
\newcommand{\E}{\operatorname{E}} %
\newcommand{\Cov}{\operatorname{Cov}} %
\newcommand{\MSE}{\operatorname{MSE}}

\DeclareMathOperator{\Nor}{N} % Normal -
\DeclareMathOperator{\Log}{Log} % Logistische Verteilung -
\newcommand{\ml}[2][1]{% % für Maximum-Likelihood-Schätzer von #1
\ifthenelse{#1 = 1}%
 {\hat{#2}_{\scriptscriptstyle{ML}}}%
 {\hat{#2}^{#1}_{\scriptscriptstyle{ML}}}% z.B. für sigmadach^2
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Environment for Aufgaben
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcounter{ka}


\newtheorem{exercise}{Problem}
\newenvironment{aufgabe}{\begin{exercise}\sf}{\end{exercise} \bigskip}

\newcommand{\footer}{ \vfill
  \mbox{}\hrulefill\\
  Upload your solutions until the 2.6.2016, 10h00 s.t. on
  Moodle. Only one solution per group. Please upload the source files
  as well as the solutions.}
%\xxivtime   Uhrzeit für letzte Änderung

\pagestyle{empty}

\renewcommand{\labelenumi}{(\alph{enumi})}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}

\renewcommand{\baselinestretch}{1}

\hrulefill\\
{\bf WiMa-Praktikum II -- Stochastic} \hspace{\fill} Summer term 2016\\
Arthur Allignol \hspace{\fill} 23.5.2016\\[-1.2ex]
\mbox{}\hrulefill\\
\newline \renewcommand{\baselinestretch}{1}
\setcounter{ka}{0} \vspace{-0.5cm}



\begin{center}
\large{{\bf Exercise Sheet 3 --- Graphics and Simple Inference}}
\end{center}



\begin{aufgabe}{{\bf Acupuncture for chronic headache in primary care:
      large, pragmatic, randomised trial}}

  Consider the data set {\tt acupuncture\_trial.xls} that contains
  data of a study on acupuncture for the treatment of chronic headache
  disorders. Information on the data and the original paper can be
  found at \url{http://www.bmj.com/content/328/7442/744.full}. The
  variables that we will consider for analysis are
  \begin{description}
  \item[{\tt id}] identification number
  \item[{\tt age}] age
  \item[{\tt sex}] gender
  \item[{\tt group}] 0: control; 1: acupuncture
  \item[{\tt pk1}, {\tt pk2}, {\tt pk5}] Severity of headache at
    visits at baseline, 6 month and 1 year, respectively. The severity
    score is assessed on a likert scale that ranges from 0: ``no
    headache'' to 5: ``Intense, incapacitating headache''. Severity of
    headache was recorded four times a day and the total summed to
    give a headache score
  \item[{\tt f1} to {\tt f5}] Frequency of headache
  \item[{\tt gen1} to {\tt gen5}] SF36 general health score. The SF36
    is a quality of life questionnaire looking at several dimensions
    of well being. The general health score is a summary quantity. 
  \item[{\tt completer}] 1 if the patient gave data until the end of
    the trial 
  \end{description}
  
  \begin{enumerate}
  \item Compare the change between baseline and 1 year of severity of
    headache, frequency of headache and SF36 general health score
    between the controls and acupuncture groups. Discuss on which test
    is appropriate for each outcome variables. 
  \item The significance level for each of the three tests is fixed at
    $\alpha = 5\%$. What is the probability of making at least 1 error
    in the 3 tests?
  \item Devise graphics that explore the evolution of the severity
    score, frequency of headaches and SF36 general health score over
    the 3 visits in each group. Conclude.
  \item Perform a graphical analysis to study whether age, gender and
    migraine influence the response to treatment, e.g., does
    acupuncture works better for migraine patients as compared to
    controls?
  \item Compare the change between baseline and visit 2 of severity of
    headache between patients who completed the trial and those who
    haven't (variable {\tt completer}). Are the conclusions drawn in
    question (a) still appropriate?
  \end{enumerate}

\end{aufgabe}

\footer


\end{document}
